---
reportdepartment: "Cardiothoracics"
title:  | 
  | Enter Title Here
  | Subheading
author: Rory Denham 
date: \today
params:
  logo: "Images/SESLHDlogo.png"
  cover: "Images/POWcover.png"
  iblue: 2b4894
  igray: d4dbde

documentclass: article
fontsize: 10
papersize: a4paper
output: 
  IReports::businessReport:
    keep_tex: TRUE
    latex_engine: xelatex
    resetStyleFiles: FALSE
header-includes: 
  - \newcommand{\logo}{`r gsub("_", "\\_", params$logo)`}
  - \newcommand{\cover}{`r gsub("_", "\\_", params$cover)`}
  - \newcommand{\iblue}{`r params$iblue`}
  - \newcommand{\igray}{`r params$igray`}
include-before:
  - \renewcommand{\contentsname}{ }
  - \renewcommand{\pagename}{Page}
---



```{r setup, include = FALSE}
# packages
library(dplyr)
library(knitr)
library(xtable)
library(skimr)
library(tidyverse)
library(magrittr)
library(lubridate)

#Debugging verbose explaination
options(tinytex.verbose = TRUE)
# settings
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r Load Data, include=FALSE}
df <- read_csv(file = "Data/qry_Sternals_and_leg_infections.csv", 
               col_types = cols(`Patient Details` = col_character(),
                                InfSite = col_factor(levels = c("Sternum", "Leg")),
                                Inf_Category = col_factor(levels = c("Follow Up", "Initial Admission")),
                                AccessInfDepth = col_factor(levels = c("Superficial", "Deep")),
                                DonorSiteDepth = col_factor(levels = c("Superficial", "Deep")),
                                Inf_Organism1 = col_guess(),
                                Gram_Positive1 = col_logical(),
                                Inf_Organism2 = col_guess(),
                                Gram_Positive2 = col_logical(),
                                Inf_Organism3 = col_guess(),
                                Gram_Positive3 = col_logical(),
                                OpDate = col_datetime(format = "%d/%m/%Y %H:%M")))


df %<>%  
  mutate(Gram_Positive_All = if_else(Gram_Positive1 == T | Gram_Positive2 == T | Gram_Positive3 == T, T, F, missing = F),
         Gram_Negative_All = if_else(Gram_Positive1 == F | Gram_Positive2 == F | Gram_Positive3 == F, T, F, missing = F)) %>% 
  mutate(Organism_Missing = if_else(Gram_Positive_All == F & Gram_Negative_All == F, T, F))

skim(df)
```


\maketitle
\tableofcontents
\addcontentsline{toc}{section}{Contents}
\clearpage

\section{Overview}

This summary report uses data derived from the local Prince of Wales Cardiothoracic Database to outline the incidence of infections from 01/01/2014 - 28/02/2019 inclusive.
\newline
Produced for presentation during a combined CTsurgery, Infectious Diseases, and Infection Control meeting on Thursday the 27th of June at 7:20am.
\newline
All statistical analysis conducted using R statistical programming language. Fully reproducible code is available upon request.
\newline\newline


Infections reviewed by Dr Fong (Advanced Trainee CTS)
\newline
Summary Statistics prepared by Rory Denham (Cardiothoracic Data Manager)


\section{Infections Summary}

\subsection{All Infections}
```{r}
plotdata1 <- df  %>%
  group_by(OpDate = floor_date(OpDate, unit = "month")) %>% 
  summarize(Infections = n(), 
            `Gram Positive` = sum(Gram_Positive_All, na.rm = T),
            `Gram Negative` = sum(Gram_Negative_All, na.rm = T),
            `Missing Organism` = sum(Organism_Missing, na.rm = T))

plot1 <- ggplot(plotdata1, aes(x = OpDate)) +
  # geom_line(aes(y = Infections, colour = "Total Infections")) +
  # geom_line(aes(y = `Gram Positive`, colour = "Gram Positive")) +
  # geom_line(aes(y = `Gram Negative`, colour = "Gram Negative")) +
  stat_smooth(aes(y = Infections, colour = "Total Infections"), span = 0.2, se = FALSE) +
  stat_smooth(aes(y = `Gram Positive`, colour = "Gram Positive"), span = 0.2, se = FALSE) +
  stat_smooth(aes(y = `Gram Negative`, colour = "Gram Negative"), span = 0.2, se = FALSE) +
  scale_x_datetime(date_breaks="4 months", date_labels="%b\n%Y", minor_breaks = "month") +
  scale_y_continuous() +
  theme_minimal() +
  theme(legend.position = "bottom")
plot1
```
\newpage

\subsection{Sternal Infection}
```{r}
plotdata1 <- df  %>%
  group_by(OpDate = floor_date(OpDate, unit = "month")) %>% 
  summarize(Infections = n(), 
            `Gram Positive` = sum(Gram_Positive_All, na.rm = T),
            `Gram Negative` = sum(Gram_Negative_All, na.rm = T),
            `Missing Organism` = sum(Organism_Missing, na.rm = T))

plot1 <- ggplot(plotdata1, aes(x = OpDate)) +
  # geom_line(aes(y = Infections, colour = "Total Infections")) +
  # geom_line(aes(y = `Gram Positive`, colour = "Gram Positive")) +
  # geom_line(aes(y = `Gram Negative`, colour = "Gram Negative")) +
  stat_smooth(aes(y = Infections, colour = "Total Infections"), span = 0.2, se = FALSE) +
  stat_smooth(aes(y = `Gram Positive`, colour = "Gram Positive"), span = 0.2, se = FALSE) +
  stat_smooth(aes(y = `Gram Negative`, colour = "Gram Negative"), span = 0.2, se = FALSE) +
  scale_x_datetime(date_breaks="4 months", date_labels="%b\n%Y", minor_breaks = "month") +
  scale_y_continuous() +
  theme_minimal() +
  theme(legend.position = "bottom")
plot1
```


\section{Major Organisms}

*10 Most common organisms over the 01/01/2014 - 28/02/2019 period.

\subsection{All Infections (Leg+Sternal)}

```{r}
All_Organisms <- df %>% 
  group_by(c(Inf_Organism1)) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(`Count`))
All_Organisms
```
\newpage

\subsection{Sternal Infections - Superficial}

```{r}
Superficial_Sternal_Organisms <- df %>% 
  filter(InfSite == "Sternum", 
         AccessInfDepth == "Superficial") %>% 
  group_by(c(Inf_Organism1)) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(`Count`))
Superficial_Sternal_Organisms
```


\subsection{Sternal Infections - Deep}

```{r}
Deep_Sternal_Organisms <- df %>% 
  filter(InfSite == "Sternum",
         AccessInfDepth == "Deep") %>% 
  group_by(c(Inf_Organism1)) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(`Count`))
Deep_Sternal_Organisms
```

\newpage

\subsection{Leg Infections - Superficial}

```{r}
Deep_Sternal_Organisms <- df %>% 
  filter(InfSite == "Leg",
         DonorSiteDepth == "Superficial") %>% 
  group_by(c(Inf_Organism1)) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(`Count`))
Deep_Sternal_Organisms
```

\subsection{Leg Infections - Deep}

```{r}
Deep_Sternal_Organisms <- df %>% 
  filter(InfSite == "Leg",
         DonorSiteDepth == "Deep") %>% 
  group_by(c(Inf_Organism1)) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(`Count`))
Deep_Sternal_Organisms
```

\newpage

\section{Superficial Leg Wound Infections Per Year}

*Note that leg wound classification as deep vs superficial began in late 2016 for the CTS database.


```{r echo=FALSE}
newdf <- df
newdf$OpDate <- as.Date(as.POSIXct(df$OpDate), tz="Australia/Sydney")
superficial_leg <- newdf %>% 
  dplyr::filter(InfSite == "Leg") %>% 
  group_by(OpYear  = floor_date(OpDate, unit = "year")) %>% 
  summarise(`Total superficial leg Infections` = as.integer(sum(if_else(DonorSiteDepth == "Superficial", 1, 0), na.rm = T)))

superficial_leg
```

```{r}
ggplot(superficial_leg, aes(x = OpYear, y = `Total superficial leg Infections`)) +
  geom_col() +
  theme_minimal()
```
\newpage

\section{Deep Leg Wound Infections Per Year}

```{r echo=FALSE}
newdf <- df
newdf$OpDate <- as.Date(as.POSIXct(df$OpDate), tz="Australia/Sydney")
deep_leg <- newdf %>% 
  dplyr::filter(InfSite == "Leg") %>% 
  group_by(OpYear  = floor_date(OpDate, unit = "year")) %>% 
  summarise(`Total Deep Leg Wound Infections` = as.integer(sum(if_else(DonorSiteDepth == "Deep", 1, 0), na.rm = T)))

deep_leg
```

```{r}
ggplot(deep_leg, aes(x = OpYear, y = `Total Deep Leg Wound Infections`)) +
  geom_col() +
  theme_minimal()
```

\newpage

\section{Superficial Sternal Wound Infections Per Year}

```{r echo=FALSE}
newdf <- df
newdf$OpDate <- as.Date(as.POSIXct(df$OpDate), tz="Australia/Sydney")
deep_leg <- newdf %>% 
  dplyr::filter(InfSite == "Sternum") %>% 
  group_by(OpYear  = floor_date(OpDate, unit = "year")) %>% 
  summarise(`Total Superficial Sternal Wound Infections` = as.integer(sum(if_else(AccessInfDepth == "Superficial", 1, 0), na.rm = T)))

deep_leg
```

```{r}
ggplot(deep_leg, aes(x = OpYear, y = `Total Superficial Sternal Wound Infections`)) +
  geom_col() +
  theme_minimal()
```

\newpage

\section{Deep Sternal Wound Infections Per Year}

```{r echo=FALSE}
newdf <- df
newdf$OpDate <- as.Date(as.POSIXct(df$OpDate), tz="Australia/Sydney")
deep_leg <- newdf %>% 
  dplyr::filter(InfSite == "Sternum") %>% 
  group_by(OpYear  = floor_date(OpDate, unit = "year")) %>% 
  summarise(`Total Deep Sternal Wound Infections` = as.integer(sum(if_else(AccessInfDepth == "Deep", 1, 0), na.rm = T)))

deep_leg
```

```{r}
ggplot(deep_leg, aes(x = OpYear, y = `Total Deep Sternal Wound Infections`)) +
  geom_col() +
  theme_minimal()
```


```{r eval=FALSE, include=FALSE}


tab <- xtable(superficial_leg, align = "|C|C|C|")

addtorow <- list()
addtorow$pos <- list()
addtorow$pos[[1]] <- -1
addtorow$command <- c("\\rowcolor{igray}")

print(tab,
      include.rownames = FALSE,
      tabular.environment = "tabularx",
      width = "\\textwidth",
      comment = FALSE,
      add.to.row = addtorow)
```

