---
reportdepartment: "Cardiothoracics"
title:  | 
  | Perfusion Data Review
  | Cardiothoracic Sugery
author: Rory Denham 
date: \today
params:
  logo: "Images/SESLHDlogo.png"
  cover: "Images/POWcover.png"
  iblue: 2b4894
  igray: d4dbde

documentclass: article
fontsize: 10
papersize: a4paper
output: 
  IReports::businessReport:
    keep_tex: TRUE
    latex_engine: xelatex
    resetStyleFiles: FALSE
header-includes: 
  - \newcommand{\logo}{`r gsub("_", "\\_", params$logo)`}
  - \newcommand{\cover}{`r gsub("_", "\\_", params$cover)`}
  - \newcommand{\iblue}{`r params$iblue`}
  - \newcommand{\igray}{`r params$igray`}
include-before:
  - \renewcommand{\contentsname}{ }
  - \renewcommand{\pagename}{Page}
---



```{r setup, include = FALSE}
# packages
library(scales)
library(kableExtra)
library(dplyr)
library(knitr)
library(xtable)
library(skimr)
library(tidyverse)
library(magrittr)
library(lubridate)
library(DataExplorer)
library(ggpubr)
library(zoo)

#Debugging verbose explaination
options(tinytex.verbose = TRUE)
# settings
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)
```


```{r Load Data, include=FALSE}
df <- read_csv(file = "Data/revised/qry_Research_CABG_Public_AL2012-2019.csv", 
               col_types = cols(OpDate = col_datetime(format = "%d/%m/%Y %H:%M:%S"),
                                Pt_Gender = readr::col_factor(levels = c(1, 2), ordered = F),
                                OpStatus = readr::col_factor(levels = c('Elective', 'Urgent'), ordered = F),
                                Adm_DOSA = col_logical(),
                                CPB_Used = col_logical(),
                                FU_Death30d = col_logical(),
                                PO_Inotrope = col_logical(),
                                PostCPB_SR = col_logical(),
                                PostCPB_DCCV = col_logical(),
                                PostCPB_Paced = col_logical(),
                                PostCPB_PaceAtrial = col_logical(),
                                PostCPB_PaceVentri = col_logical()
                                ))


df$Pt_Gender <- factor(df$Pt_Gender, levels = 1:2, labels = c('Male', 'Female'))

#drop useless columns
df_clean = subset(df, select = -c(`Hospital`, `PreOp_MI`, `PreOp_Resus`, `PreOp_Arrhythmia`, `Prev_CardiacProcedure`, `Prev_CardiacSurgery`, `OpDescription`, `Death_30Days_OR_InHosp`, `FU_Death`))
```



\maketitle
\tableofcontents
\addcontentsline{toc}{section}{Contents}
\clearpage

\section{Overview}

This summary report uses data derived from the local Prince of Wales Cardiothoracic Database to investigate troponin changes in Isolated Coronary Artery Bypass Grafts (CABG) cases from 01/01/2012 - 31/12/2018 inclusive.
\newline
\newline
Produced for CTS Perfusion Dept. Prince of Wales Hospital.
\newline
\newline
All statistical analysis conducted using R statistical programming language. 
\newline
Fully reproducible code is available upon request.
\newline
\newline
Summary Statistics prepared by Rory Denham (Cardiothoracic Data Manager)
\newline
\newline
\newline

\Large Data Parameters: \newline
\normalsize
Date Range: 01/01/2012 - 31/12/2018 inclusive \newline
Exclusions: \newline
- Hospital = Prince_of_Wales_Public \newline
- PreOp_MI = FALSE \newline
- PreOp_Resus = FALSE \newline
- PreOp_Arrhythmia = FALSE \newline
- Prev_CardiacSurgery = FALSE \newline
- OpDescription = CABG (no OPCAB) \newline


\section{Data Overview}

\subsection{exclusions}

After a preliminary look at the dataset the following cleaning processes were applied to the raw dataset:

#clean continuous variables
\newline
df_clean %<>% mutate(Path_MaxCreat = if_else(Path_MaxCreat > 10, NA_real_, Path_MaxCreat)) #remove values greater than 10
\newline
df_clean %<>% mutate(Path_DischCreat = if_else(Path_DischCreat > 10, NA_real_, Path_DischCreat)) #remove values greater than 10
\newline
df_clean %<>% mutate(Path_DischHb = if_else(Path_DischHb < 20, NA_real_, Path_DischHb)) #remove values less than 20
\newline
df_clean %<>% mutate(CPB_LowTemp = if_else(CPB_LowTemp > 90, NA_real_, CPB_LowTemp))   #remove values with CPB entries over 90
\newline
df_clean %<>% mutate(VentHrs_Total = if_else(VentHrs_Total > 24 * 7 * 4 *3, NA_real_, VentHrs_Total)) #where intubation is greater than 3 months (likely an input error) relabel the value as missing
\newline
df_clean %<>% mutate(VentHrs_IntraOp = if_else(VentHrs_IntraOp > 200, NA_real_, VentHrs_IntraOp))
\newline
\newline

df_clean <- df_clean[!is.na(df_clean$OpDate),]

```{r}
#clean continuous variables
df_clean %<>% mutate(Path_MaxCreat = if_else(Path_MaxCreat > 10, NA_real_, Path_MaxCreat)) #remove values greater than 10
df_clean %<>% mutate(Path_DischCreat = if_else(Path_DischCreat > 10, NA_real_, Path_DischCreat)) #remove values greater than 10
df_clean %<>% mutate(Path_DischHb = if_else(Path_DischHb < 20, NA_real_, Path_DischHb)) #remove values less than 20
df_clean %<>% mutate(CPB_LowTemp = if_else(CPB_LowTemp > 90, NA_real_, CPB_LowTemp))   #remove values with CPB entries over 90
df_clean %<>% mutate(VentHrs_Total = if_else(VentHrs_Total > 24 * 7 * 4 *3, NA_real_, VentHrs_Total)) #where intubation is greater than 3 months (likely an input error) relabel the value as missing
df_clean %<>% mutate(VentHrs_IntraOp = if_else(VentHrs_IntraOp > 200, NA_real_, VentHrs_IntraOp))

df_clean <- df_clean[!is.na(df_clean$OpDate),]
```

\newpage

\subsection{Summary}


All Data
\newline
\newline

```{r echo=FALSE, results='asis'}
skim_with(numeric = list(iqr = NULL, p25 = NULL, p50= NULL, p75 = NULL, hist = NULL),
          integer = list(iqr = NULL, p25 = NULL, p50= NULL, p75 = NULL, hist = NULL))

skim(df_clean) %>% kable(format = "latex", booktabs = T)
```

\subsection{Numeric Variable Distributions}

```{r}
plot_histogram(df_clean)
```
\newpage
\section{Pathology Investigations}

\subsection{Pathology levels by Year}

```{r}
plotdata <- df_clean %>% 
  select(OpDate, Path_MaxTrop, Path_PreOpHb, Path_PreOpCreat, Path_DischHb, Path_DischCreat, Path_MaxCreat) %>% 
  mutate(OpYear = year(OpDate))

iteration_string <- c("Path_MaxTrop", "Path_PreOpHb", "Path_DischHb", "Path_PreOpCreat", "Path_DischCreat", "Path_MaxCreat")
p = list()
for(y_variable in iteration_string){
      p <- ggboxplot(plotdata, 
                     title = y_variable,
                   x = "OpYear",
                   y = y_variable,
                   color = "OpYear",
                   palette = "lancet",
                   add = "jitter",
                   outlier.shape = NA,
                   ylab = y_variable)
  print(p)
}
```

\subsection{Troponin Box Plot - Comparison with all Variables}

Plot troponin as ordinal groups

```{r}
plot_boxplot(df_clean, 
             by = "Path_MaxTrop", 
             ncol = 2)
```

\section{CPB Measures}

\subsection{CPB Measures by Year}

```{r}
plotdata <- df_clean %>% 
  select(OpDate, CPB_BMI, CPB_PerfTime, CPB_XClampTime, CPB_LowTemp) %>% 
  mutate(OpYear = year(OpDate))

iteration_string <- c('CPB_BMI', 'CPB_PerfTime', 'CPB_XClampTime', 'CPB_LowTemp')
p = list()
for(y_variable in iteration_string){
      p <- ggboxplot(plotdata,
                     title = y_variable,
                   x = "OpYear",
                   y = y_variable,
                   color = "OpYear",
                   palette = "lancet",
                   outlier.shape = NA,
                   add = "jitter",
                   ylab = y_variable)
  print(p)
}
```
\subsection{Operation Count Per Year}

```{r}
#cases per year?
df_clean %>% group_by(year(OpDate)) %>% 
  summarise(Count = n())
```

\section{Length of Stay}

\subsection{Length of Stay by Year}

```{r}
plotdata <- df_clean %>% 
  select(OpDate, Op_PreOpLOS, Op_PostOpLOS, Op_TotalLOS) %>% 
  mutate(OpYear = year(OpDate))

iteration_string <- c('Op_PreOpLOS', 'Op_PostOpLOS', 'Op_TotalLOS')

for(y_variable in iteration_string){
      p <- ggboxplot(plotdata,
                   title = y_variable,
                   x = "OpYear",
                   y = y_variable,
                   color = "OpYear",
                   palette = "lancet",
                   add = "jitter",
                   outlier.shape = NA,
                   ylab = y_variable)
  print(p)
}
```

\subsection{Length of Stay by Operation Status}

*Dashed line represents the group mean

```{r}

plotdata <- df_clean %>% 
  select(OpDate, OpStatus, Op_PreOpLOS, Op_PostOpLOS, Op_TotalLOS) %>% 
  mutate(OpYear = year(OpDate))

iteration_string <- c('Op_PreOpLOS', 'Op_PostOpLOS', 'Op_TotalLOS')

for(x_variable in iteration_string){
      p <- gghistogram(plotdata,
                      x = x_variable,
                      title = x_variable,
                      add = "mean", 
                      rug = TRUE,
                      color = "OpStatus", 
                      fill = "OpStatus",
                      palette = c("#00AFBB", "#E7B800"))
  print(p)
}
```

\subsection{Grouped Barcharts Over Time}


```{r}

variable_list <- c(quo(Pt_Gender), quo(OpStatus))

for(factor_variable in variable_list) {
  #we want to build a dataset with 4 columns in order to plot our data correctly.
  #x axis ticks are plotted as year/month
  #two columns displayed for each x axis tick (eg. sex - male and female). This is our variable of interest.
  #lastly we are facetting by year so that the graph does not become too cluttered when plotting multiple years.
  factor_variable <- enquo(factor_variable)
  
  plot_df <- df_clean %>% 
  mutate(year_month = floor_date(OpDate, unit = "months")) # generate x axis column
  
  plot_df <- plot_df %>% 
  group_by(year_month, !! factor_variable, .drop = FALSE) %>% 
  tally() #summarise a count for each factor within our variable of interest, for each year/month.

  plot_df %<>% mutate(year = year(year_month)) #create year variable after summarising (if done before we lose the)

  plot <- ggplot(plot_df, aes(year_month, n, fill = !! factor_variable)) + 
    geom_bar(position= position_dodge2(width = 0.9), stat="identity") + 
    facet_wrap(. ~ year, scales = "free", ncol = 2) +
    scale_x_datetime(labels = date_format("%b"), breaks='1 month')
    
  print(plot)
}
```

```{r}

plot_df <- df_clean %>% 
  mutate(year_month = floor_date(OpDate, unit = "months")) # generate x axis column  

plot_df <- plot_df %>% 
  group_by(year_month, .drop = FALSE) %>% 
  summarise(Mean_HCT = mean(Path_CPBHct)) #summarise a count for each factor within our variable of interest, for each year/month.
  
plot_df %<>% mutate(year = year(year_month)) #create year variable after summarising (if done before we lose the)


plot <- ggplot(plot_df, aes(year_month, Mean_HCT)) + 
  geom_bar(position= position_dodge2(width = 0.9), stat="identity") +
  facet_wrap(. ~ year, scales = "free", ncol = 2) +
  scale_x_datetime(labels = date_format("%b"), breaks='1 month')

print(plot)
```

\subsection{Transfusions Over Time}

```{r}
plot_df <- df_clean %>% 
  mutate(year_month = floor_date(OpDate, unit = "months")) # generate x axis column  

plot_df <- plot_df %>% 
  group_by(year_month, .drop = FALSE) %>% 
  summarise(Mean_CPB_RBC = mean(Bld_CPB_RBC),
            Mean_CPB_FFP = mean(Bld_CPB_FFP),
            Mean_CPB_PLT = mean(Bld_CPB_Plt)) #summarise a count for each factor within our variable of interest, for each year/month.
  
plot_df %<>% mutate(year = year(year_month)) #create year variable after summarising (if done before we lose the)

variable_list <- c(quo(Mean_CPB_RBC), quo(Mean_CPB_PLT), quo(Mean_CPB_PLT))

for(factor_variable in variable_list) {
  #we want to build a dataset with 4 columns in order to plot our data correctly.
  #x axis ticks are plotted as year/month
  #two columns displayed for each x axis tick (eg. sex - male and female). This is our variable of interest.
  #lastly we are facetting by year so that the graph does not become too cluttered when plotting multiple years.
  factor_variable <- enquo(factor_variable)

  plot <- ggplot(plot_df, aes(year_month, !!factor_variable)) + 
    geom_bar(position= position_dodge2(width = 0.9), stat="identity") +
    facet_wrap(. ~ year, scales = "free", ncol = 2) +
    scale_x_datetime(labels = date_format("%b"), breaks='1 month')

  print(plot)
}
```

Transfusions by Year

```{r}
df_clean %>% 
  group_by(year(OpDate)) %>% 
  summarise(Units_RBC = sum(Bld_CPB_RBC),
            Units_FFP = sum(Bld_CPB_FFP, na.rm = T),
            Units_Plt = sum(Bld_CPB_Plt, na.rm = T))
```


